S2 Appendix

### 2021-02-01 src/s2.Rmd
visualisations

---
title: "S2 Appendix"
author: ""
output: pdf_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_resubmission")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(car)
library(lmtest)
library(plm)
library(stargazer)
library(knitr)

un <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_un_regressions.csv")
us <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_usa_regressions.csv")
uk <- read_csv("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/data_for_uk_regressions.csv")
```

# First stage regressions for the 2SLS estimations

## Global model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

net.2sls.s1 <- lm(net100hab ~  fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

bb.2sls.s1 <- lm(bb100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

fixed.2sls.s1 <- lm(fixed100hab ~ fem.lab + log(pop.dens) + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

mobile.2sls.s1 <- lm(mobile100hab ~ fem.lab + log(pop.dens)  + gov.exp + trade + non.agri + gdp.growth + log(gdppc) +  log(pop) 
                      + as.factor(year) + as.factor(country), 
               data = un, weights = 1/(se^2))

net.2sls.s1.rse <- coeftest(net.2sls.s1, function(x) vcovHC(x, type="HC0"))

bb.2sls.s1.rse <- coeftest(bb.2sls.s1, function(x) vcovHC(x, type="HC0"))

fixed.2sls.s1.rse <- coeftest(fixed.2sls.s1, function(x) vcovHC(x, type="HC0"))

mobile.2sls.s1.rse <- coeftest(mobile.2sls.s1, function(x) vcovHC(x, type="HC0"))

stargazer(net.2sls.s1, bb.2sls.s1, fixed.2sls.s1, mobile.2sls.s1, type = "latex",
          se = list(net.2sls.s1.rse[,"Std. Error"], bb.2sls.s1.rse[,"Std. Error"],
                    fixed.2sls.s1.rse[,"Std. Error"], mobile.2sls.s1.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 5 \\label{2sls.global.s1}",
          font.size = "footnotesize",
          column.sep.width = "1pt",
          # dep.var.labels=c("Internet users per 100 hab. (log)",
          #                 "Broadband users per 100 hab. (log)",
          #                 "Mobile phone users per 100 hab. (log)",
          #                 "Fixed phone users per 100 hab. (log)"),
          column.labels=c('Internet users',
                           "Broadband users",
                           "Mobile phone users",
                           "Fixed phone users"),
          dep.var.labels.include = FALSE,
          #column.separate = c(1,1,1,1),
          covariate.labels=c("Female labour force (\\%)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(
            # c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
            #                  round(summ.net.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
            #                c(rownames(summ.net.2sls$diagnostics)[2],
            #                  round(summ.net.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
            #                c("P-value",
            #                  round(summ.net.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
            #                  round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust Std. Errors in parenthesis"))

net.test <- car::linearHypothesis(net.2sls.s1, c("fem.lab=0"))
bb.test <- car::linearHypothesis(bb.2sls.s1, c("fem.lab=0"))
mob.test <- car::linearHypothesis(fixed.2sls.s1, c("fem.lab=0"))
fixed.test <- car::linearHypothesis(mobile.2sls.s1, c("fem.lab=0"))

tests <- data.frame("F" = c(net.test$F[2], bb.test$F[2], mob.test$F[2], fixed.test$F[2]),
                    "p" = c(net.test$`Pr(>F)`[2], bb.test$`Pr(>F)`[2], mob.test$`Pr(>F)`[2], fixed.test$`Pr(>F)`[2]),
                    row.names = c("internet users", "broadband users", "mobile phone users", "fixed telephony users"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IV (female labour force (\\%)) from first stage regressions for Table 5 \\label{2sls.global.s1.f}",
          header = FALSE
)
```

\newpage

## US model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(hh.net.share ~ bachelors2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

model_iv2.s1 <- lm(hh.net.share ~ bachelors2005 + commuting.minutes2005 + 
                     log(pop2013) + un13 + white2013share + 
                     log(income2013) + I(pop2013/area_h) + service2013 + commuting.minutes2013, 
                   data = us)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0")) # rob = rse, naming inconsistency
rob.model_iv2.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1, 
          type = "latex",
          title = "First stage regressions for Table 8 \\label{2sls.us.s1}",
          header = FALSE,
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]), 
          dep.var.labels=c("\\% of households w. internet 2013"),
          covariate.labels=c("bachelors degree per hab. 2005",
                             "commute in minutes 2005",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2012"),
          single.row = FALSE,
          df = FALSE, 
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Error in parenthesis"))

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("bachelors2005=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("bachelors2005=0","commuting.minutes2005=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("bachelors degree per hab., 2005", "the above + commute in minutes, 2005"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)
```

\newpage

## UK model

```{r, echo=FALSE, warning=FALSE, results='asis', message=FALSE}

model_iv1.s1 <- lm(log(download2011) ~ uni.freq +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

model_iv2.s1 <- lm(log(download2011) ~ uni.freq + n.tests2011 +
                     log(pop2011) + I(n.tests2011/pop2011) + unemployment.share + 
                     british.pop.share +  pop.dens + work.home.share + services.share
                     , data = uk, subset = n.tests2011>30)

rob.model_iv1.s1 <- coeftest(model_iv1.s1, function(x) vcovHC(x, type="HC0"))
rob.model_iv2.s1 <- coeftest(model_iv2.s1, function(x) vcovHC(x, type="HC0"))

stargazer(model_iv1.s1, model_iv2.s1,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "First stage regressions for Table 11 \\label{2sls.us.s1}",
          font.size = "small",
          se = list(rob.model_iv1.s1[,"Std. Error"], rob.model_iv2.s1[,"Std. Error"]),
          dep.var.labels=c("download speed, 2011 (log)"),
          covariate.labels=c("N. of universities",
                             "N. of broadband tests, 2011",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), 
          notes = c("Robust Std. Errors"))

# N. of universities",
# N. of universities, N. of broadband tests, 2011"),

model_iv1.s1.test <- car::linearHypothesis(model_iv1.s1, c("uni.freq=0"))
model_iv2.s1.test <- car::linearHypothesis(model_iv2.s1, c("uni.freq=0", "n.tests2011=0"))

tests <- data.frame("F" = c(model_iv1.s1.test$F[2], model_iv2.s1.test$F[2]),
                    "p" = c(model_iv1.s1.test$`Pr(>F)`[2], model_iv2.s1.test$`Pr(>F)`[2]),
                    row.names = c("N. of universities", "the above + N. of broadband tests, 2011"))

stargazer(tests, 
          summary = F, 
          type = "latex",
          digits = 3,
          title = "F tests for IVs from first stage regressions for Table 8 \\label{2sls.us.s1.f}",
          header = FALSE
)

>>>>

---
title: "S2 Appendix"
author: ""
output: html_document
header-includes:
- \usepackage{multirow}
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../paper/2_final_submission")
  })
---

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}
library(rmarkdown)
library(tidyverse)
library(plotly)
library(rnaturalearth)
library(rnaturalearthdata)
library(gridExtra)
library(cowplot)
library(rprojroot)

path <- find_rstudio_root_file()
```

# Pareto exponents

The below maps plot the Pareto exponents for all the countries included in our data 
set for the years 2000, 2005, 2010 and 2010. The quantiles of the yearly exponent 
distribution are used for the maps with the first quantile always representing 
the countries with the of the smallest Pareto exponent.

```{r, echo=FALSE, warning=FALSE, fig.width = 8}

# load data
path.data <- paste0(path, "/data/data_inter/UN.RData")
load(path.data)

# a zipf table for the appendix, which includes the Zipf coef. per year and country
zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Countries` = `as.factor(country)`)

# a world `sf` to join with Zipf
world <- ne_countries(scale = "medium", returnclass = "sf")
world <- merge(world, zipf.table, by.x = 'iso_a3', by.y = 'Countries', all = T)

# yearly maps
map2000 <- world %>%
  mutate(eq.n2000 = cut_number(`2000`, 
                               n = 4,
                               labels = c("1st quantile", "2nd quantile",
                                          "3rd quantile", "4th quantile"))) %>%
  ggplot() +
  geom_sf(aes(fill = eq.n2000)) + 
  ggtitle("2000") +
  theme_minimal() + 
  theme(legend.position = "top",
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

legend <- get_legend(map2000)
map2000 <- map2000 + theme(legend.position="none")

map2005 <- world %>%
  mutate(eq.n2005 = cut_number(`2005`, 
                               n = 4,
                               labels = c("1st quantile", "2nd quantile",
                                          "3rd quantile", "4th quantile"))) %>%
  ggplot() +
  geom_sf(aes(fill = eq.n2005), show.legend = F) + 
  ggtitle("2005") +
  theme_minimal() + 
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

map2010 <- world %>%
  mutate(eq.n2010 = cut_number(`2010`, 
                               n = 4,
                               labels = c("1st quantile", "2nd quantile",
                                          "3rd quantile", "4th quantile"))) %>%
  ggplot() +
  geom_sf(aes(fill = eq.n2010), show.legend = F) + 
  ggtitle("2010") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

map2015 <- world %>%
  mutate(eq.n2015 = cut_number(`2015`, 
                               n = 4,
                               labels = c("1st quantile", "2nd quantile",
                                          "3rd quantile", "4th quantile"))) %>%
  ggplot() +
  geom_sf(aes(fill = eq.n2015), show.legend = F) + 
  ggtitle("2015") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

lay <- rbind(c(1,2),
             c(3,4),
             c(5,5))
grid.arrange(map2000, map2005, map2010, map2015, legend,
            nrow=3, 
            heights=c(2.3, 2.3, 0.3),
            layout_matrix = lay)
```

The Pareto exponents are plotted below as an interactive line graph.

```{r, echo=FALSE, warning=FALSE, fig.width = 8}

zipf.line <- zipf.table %>% 
  pivot_longer(!Countries, names_to = "year", values_to = "zipf") %>% 
  ggplot(aes(x=year, y=zipf, group = Countries, colour = Countries),
         show.legend = F) + 
  geom_line() + guides(colour=FALSE) + xlab("Year") +
  ylab("Pareto exponent")

ggplotly(zipf.line,
         show.legend=F)
